### 2015-10-31 pde/pdetest1.tex
ode and paper

>>>>

\section{The Cauchy Problem}
\noindent \\\\
\indent Up until now, we have always been given our PDE's with boundary conditions and initial values; but imagine that we did not have a boundary? The traditional method that we have been using would fail at this point. Since we would not be able to find any eigenvalues, we would not be able to form a series solution by using the \textit{Separation of Variables} technique. Let's consider the familiar heat equation:
\xi_{t} - k\xi_{xx} = 0\\
\xi(x,0) = f(x)
Through much careful derivation and theory, we can say that the general solution for this problem will be
\xi(x,t) = \frac{1}{\sqrt{4 \pi k t}}\int_{-\infty}^{\infty}e^{-\frac{(x-y)^{2}}{4kt}}f(y)dy.
The question now should be ``How can we solve this?'' The first step we will take is letting
u = \frac{y-x}{\sqrt{4kt}}.
We are going to perform a \textit{u substitution} to try and put this integral into a nicer form. This yields
y &= x + u\sqrt{4kt}\\
dy &= \sqrt{4kt}\ du.
This changes our original problem into
\xi(x,t) = \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}f(x+u\sqrt{4kt})du.
Let's look at an example\\\\
\mu_{t} - k\mu_{xx} = 0\\
\mu(x,0) = \cos{x}
\indent \textbf{\textit{Solution:}} From above, we know what our general solution will look like. All we need to do is substitute in our initial condition. We have the following:
\mu(x,t) &= \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}\cos{(x + u\sqrt{4kt})}du\\
&= \frac{1}{\sqrt{\pi}}\int_{-\infty}^{\infty}e^{-u^{2}}\left[\cos{(x)}\cos{(u\sqrt{4kt})} - \sin{(x)}\sin{(u\sqrt{4kt})}\right]du\\
&=  \frac{2}{\sqrt{\pi}}\cos{(x)}\int_{0}^{\infty}e^{-u^{2}}\cos{(u\sqrt{4kt})}du.
You may be wondering what happened to the $\sin$ terms. Recall that $\sin$ is an odd function, so it will ultimately be $0$ on the interval $[-\infty,\infty]$; because of this fact, we can remove it now. We also use the face that the $\cos$ is even, so it is okay to halve the interval and multiply by two. Now we need to evaluate this integral. We will substitute $\alpha = \sqrt{4kt}$, and assign a function, $\Upsilon(\alpha)$, to the integral. This gives us the following system:
\Upsilon(\alpha) = \int_{0}^{\infty}e^{-u^{2}}\cos{(\alpha u)}du\\
\Upsilon(0) = \int_{0}^{\infty}e^{-u^{2}}du\\
Believe it or not, we actually know the answer to $\Upsilon(0)$. To find it, we can assign a function to the integral in a similar fashion as we have already done. Calling it $\Psi$, we have
\Psi &= \int_{0}^{\infty}e^{-u^{2}}du\\
\Psi^{2} &= \int_{0}^{\infty}e^{-u^{2}}du \int_{0}^{\infty}e^{-v^{2}}dv\\
\Psi^{2} &= \int_{0}^{\infty}\int_{0}^{\infty}e^{-u^{2}-v^{2}}dudv\\
\Psi^{2} &= \int_{0}^{\pi/2}\int_{0}^{\infty}re^{-r^{2}}drd\theta\\
\Psi^{2} &= \frac{\pi}{2}\left[-\frac{1}{2}e^{-r^{2}}\right]\Bigg |_{r=0}^{r=\infty}\\
\Psi^{2} &= \frac{\pi}{4}\\
\Psi &= \frac{\sqrt{\pi}}{2}.
So we know that $\Upsilon(0) = \frac{\sqrt{\pi}}{2}$. Plugging this into our system gives us
\Upsilon(\alpha) = \int_{0}^{\infty}e^{-u^{2}}\cos{(\alpha u)}du\\
\Upsilon(0) = \frac{\sqrt{\pi}}{2}\\
Now, we will take the derivative of $\Upsilon$. This gives us
\Upsilon'(\alpha) = -\int_{0}^{\infty}ue^{-u^{2}}\sin{(\alpha u)}du.
Integration by parts gives us
\frac{1}{2}e^{-u^{2}}\sin{(\alpha u)}\Big |_{0}^{\infty} - \alpha\int_{0}^{\infty}\frac{1}{2}e^{-u^{2}}\cos{(\alpha u)}du.
The value from $0$ to $\infty$ is zero, since the $\sin$ function odd. This leaves us with following relationship:
\Upsilon'(\alpha) = -\frac{\alpha}{2}\Upsilon(\alpha) \alpha.
We can solve this as an ODE. It's separable, so we get
\frac{d\Upsilon}{\Upsilon} &= -\frac{\alpha}{2}d\alpha\\
\ln{(\Upsilon)} &= -\frac{\alpha^{2}}{4} + c_{1}\\
\Upsilon &= c_{2}e^{-\frac{\alpha^{2}}{4}}.
Applying the initial condition gives us
\Upsilon = \frac{\sqrt{\pi}}{2}e^{-\frac{\alpha^{2}}{4}}.
Substituting back in gives us a final answer of
\mu(x,t) &= \frac{2}{\sqrt{\pi}}\cos{x}\Upsilon(\sqrt{4kt})\\
&= \cos{x}e^{-kt},
and we are done.
